---
title: "Psychrometric calculations"
author: "Ian Lyttle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

This vignette shows how to make psychrometric calculations using the `humidr` package.

Let's say you have a weather station at a given elevation:

```{r}
elevation <- 265 # elevation (m)
```

Let's say you have an set of weather observations at this station:

```{r}
t_db <- 16.7 # dry-bulb temperauture (°C)
rh <- 0.96   # relative humidity (1)
p_sl <- 101380.0 # sea-level pressure (Pa)
```

The first thing we need to do is to estimate the atmospheric pressure at the station, given the sea-level pressure reported at the weather station.

```{r}
p_atm <- p_atm_z(z = elevation, p_sl = p_sl) # atmospheric pressure (Pa)
p_atm
```

Now, we can find the humidity ratio (mass of water vapor/mass of dry air), the enthalpy of the moist air (kJ/kg), and the wet-bulb temperature (°C):

```{r}
w <- w_t_rh(t = t_db, rh = rh, p_atm = p_atm)  # humidity ratio (1)
h_moist <- h_t_w(t = t_db, w = w)              # enthalpy (kJ/kg) 
t_wb <- t_wb(t = t_db, rh = rh, p_atm = p_atm) # wet-bulb temperature (°C)

w
h_moist
t_wb
```

## Vectorized calculation

Given a series of observations:

```{r}
library(dplyr)

wx_today <- data.frame(
  t_db = c(25.0, 25.0, 22.8, 21.1, 16.7),
  rh = c(0.54, 0.54, 0.62, 0.68, 0.96),
  p_sl = c(101310.0, 101260.0, 101240.0, 101310.0, 101380.0)
)

wx_today <- 
  wx_today %>%
  mutate(
    w = w_t_rh(t_db, rh, p_atm),
    h_moist = h_t_w(t_db, w),
    t_wb = t_wb(t_db, rh, p_atm)
  )
```

```{r results='asis'}
knitr::kable(wx_today)
```

## Reference 

Chapter 6, 2005 ASHRAE Handbook -- Fundamentals (SI)

